---
title: "dataGenerato.R"
output: html_document
date: "2023-01-12"
---
```{r}
install.packages("dplyr", "ggplot2")
library(ggplot2)
```

Script #1 (dataGenerato.R):

1. Vector with a set of 100 species names, randomly selected from a vector of species 
```{r}
Species<-c("Panthera leo","Panthera pardus", "Panthera onca", "Panthera tigris", "Panthera uncia")
SpeciesList<-sample(Species, size = 100, replace = T)
SpeciesList
```

2. A vector of 100 values for limb width, randomly selected from a vector of values
```{r}
widthvalues<-c(8:18)
Limb_width<-sample(widthvalues, size=100, replace =T)
Limb_width
```

3. A vector of 100 values for limb length, randomly selected from a vector of values
```{r}
lengthvalues<-c(15:30)
Limb_length<-sample(lengthvalues, size = 100, replace = T)
Limb_length
```

4. A vector of 100 values for observer, randomly selected from a vector of names
```{r}
names<-c("Erica R", "Riley C", "Maria D")
Observer<-sample(names, size = 100, replace = T)
Observer
```

5. Dataframe that combines the Species type, limb width, limb length, and the observer
```{r}
Measurements<-data.frame(SpeciesList, Limb_width, Limb_length, Observer)
Measurements
```

Save as csv
```{r}
write.csv(Measurements, "Measurements.csv", row.names=TRUE)
```

